
# Analysis of earthquake activity in 2014 around Darmstadt, Germany

## Synopysis

In May and June 2014 several earthquakes struck the region east of Darmstadt (Germany). In this report we will use the earthquake data provided by the [Bundesanstalt für Geowissenschaften und Rohstoffe (BGR)][BGR] and the `ggmap` package to create different maps. 

_Note:_ This analysis was inspired by [Bob's (Rudis) blog post][bob] about  the series of earthquakes which hit California in March 2014.

## Loading packages

```{r message=FALSE}

library(ggplot2)
library(ggmap)
library(dplyr)
library(httr)
library(XML)
library(stringr)
library(lubridate)

```

## Downloading and processing data

We will download the data using Hadley's `httr` package and include all earthquake records from 2014 for Germany.

Actually, the `GET` call below is not necessary. You can download the data directly using `htmlParse` from the `XML` package. However, I just wanted to try it out. 

```{r}

page <- GET('http://www.seismologie.bgr.de/cgi-bin/eisy_test_iframe.pl?begin_date=20140101&end_date=20140715&min_lat=&max_lat=&min_lon=&max_lon=&min_mag=&max_mag=&')
page_content <- content(page, 'text')

parsed_content <- xmlRoot(htmlParse(page_content))

```

Let's take a look at the downloaded data: 

```{r}

parsed_content

```

The relevant earthquake records are included and surrounded by `pre` tags. We will fetch the records using a `XPath` expression and store them in a data frame. Based on the fact that they are stored in a fixed length format we can use a series of `str_sub` calls to extract the relevant parts one by one. 


```{r}

doc <- xpathSApply(parsed_content, '//pre', xmlValue)
head(doc)

data <- doc[-1]

date <- str_sub(data, 1, 10) 
time <- str_sub(data, 11, 17) %>% str_trim 
lat <- str_sub(data, 18, 25) %>% str_trim 
long <- str_sub(data, 26, 32) %>% str_trim 
mag <- str_sub(data, 33, 38) %>% str_trim 
region <- str_sub(data, 39, str_length(data)) %>% str_trim

preprocessed_data <- data.frame(date, time, lat, long, mag, region, 
                             stringsAsFactors = FALSE)
```

How does the pre-processed data look like?

```{r}

head(preprocessed_data)

```

Now, we are only interested in earthquakes which struck the region of Darmstadt. In addition, we will transform some of the character variables into numeric ones. 

```{r}

# Only include earthquakes around Darmstadt and change order so that older
# earthquakes will get plotted first in case they need to get plotted on top of 
# each other
processed_data <- filter(preprocessed_data, grepl('.*Darmstadt.*', region)) %>%
  transform(
    date  = dmy(date),
    lat   = as.numeric(lat),
    long  = as.numeric(long),
    mag   = as.numeric(mag)
    ) %>%
  arrange(date)

```

Our processed data looks as follows and can directly be used for plotting in the next section:

```{r}

processed_data 

```


## Plotting data

We’ll use `ggmap` to display the center of the reported earthquakes on a map of Darmstadt.


In our first attempt we will plot a toner map provided by [Stamen][stamen]




```{r}

# Get toner map of Darmstadt from Stamen
da_toner_map <- get_map('Darmstadt', maptype = 'toner', source = 'stamen')

# Create toner plot object
da_toner_plot <- ggmap(da_toner_map) +
  geom_point(aes(long, lat, size = mag, fill = mag), processed_data, shape = 21,
             alpha = 0.7) +
  ggtitle("2014: Recent earthquakes in Darmstadt") +
  scale_size_continuous(limits = c(2,4)) +
  scale_fill_gradientn(colours = rev(terrain.colors(5)), limits=c(2,4), 
                       guide = guide_legend()) +
  labs(size = 'Magnitude', fill = 'Magnitude') +
  theme(axis.title = element_blank())

```


```{r}

da_toner_plot

```

In our next plot we will use a terrain map provided by Google.



[bob]: http://rud.is/b/2014/04/01/mapping-the-march-2014-california-earthquake-with-ggmap/

[BGR]: http://www.bgr.bund.de/DE/Themen/Erdbeben-Gefaehrdungsanalysen/Seismologie/Seismologie/Erdbebenauswertung/D_seit_1968/d_1968_node.html

[stamen]: http://http://maps.stamen.com/#toner/
